The Challenges of Teaching Technical Writing

### 2019-03-13 Topic_categorization.Rmd
update

#Topic Categorization and Modeling Methods
Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already, or do they need to be established? What are the main topics

@@

LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion. LSA also deals effectively with sparse, ambiguous, and contradictory data.

>>>>

Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already, or do they need to be established?

@@

LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.). This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion. LSA also deals effectively with sparse, ambiguous, and contradictory data.

### 2019-03-13 background.Rmd
update

Undergraduate introductory STEM courses often include technical writing training. Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.

Adding to the learning challenge, faculty may not be the ones helping students develop writing skills. If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.

@@

*  Embody best practices from 40+ years of research on "writing in the disciplines;"
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
*  Can be implemented at scale by less experienced teaching assistants.

####\  
##Context & Prior Work
From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:

@@

Our bins-based report grading protocol reinforces this feedback strategy.

This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What is needed is a more direct measure of TA effort.

@@

We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure. This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized. Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)

@@

>>>>

Undergraduate introductory STEM courses often include technical writing training. Too often it is treated as an incidental skill that each student develops "along the way." At other times writing instruction is haphazardly designed, implemented, and executed. Adding to the learning challenge, faculty may not be the ones helping students develop writing skills. If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.

The essential first step towards improving the current situation is widespread adoption of evidence-based writing training. Equally important is reducing the practical barriers that keep faculty from routinely including scientific writing in post-secondary STEM classes.

@@

*  Embody best practices from 40+ years of research on "writing in the disciplines";
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
*  Can be implemented at scale by less experienced teaching assistants.

####\  
##Context & Prior Work
From prior studies (*see the Literature Review for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@Bazerman1994; @Bazerman2006; @Breidenbach2006; @Gottschalk2003; @Graves2011; @Kiefer2002; @NCTENWP; @Perelman2009; @Ruegg2015; @Szymanski2014;@Trupiano2006; @Underwood2006; @Anson2001; @CWPA2011; @Harris1979; @Haswell2006a; @Haswell2006b; @Reynolds2009; @Szymanski2014; @Swilky1991]. To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:

@@

Our bins-based report grading protocol [@Nilson2014] reinforces this feedback strategy.

This holistic, coaching-oriented approach is well-supported by prior research and is a recommended best practice by (among others) the Writing Across the Curriculum Clearinghouse. Yet we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What is needed is a more direct measure of TA effort.

@@

We next used a subset of 110 randomly selected comments to build an initial qualitative codebook [@Elliott2018; @Saldana2015] for classifying comments based on subject and structure. This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized. Using the final codebook, we then sorted all 11,000 comments into separate categories.

@@

####\  
***
*Work Cited*

### 2019-03-13 build.Rmd
update

####Embedding the Naive Bayes Classifier in an Ensemble Model  
*(Planning stage)*
The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category. This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\  
***

>>>>

####Embedding the Naive Bayes Classifier in an Ensemble Model  
*(Planning stage)*
The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category. This page outlines a draft plan for a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

### 2019-03-13 index.Rmd
update

In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. Using an iteratively generated codebook, each of the comments was separately classified in two categories:

* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?

#### \   
Counts of coded TA comments were tabulated according to subject and structure, and frequencies for each subject/structure pair calculated as a fraction of all comments. This 2-way contingency table provided us with a direct estimate of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.

#### \   
##Specific Goals
Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data. This automated classifier will used within the larger NSF-funded research project to:

*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
*  Monitor how the types and focus of TAs comments on student writing change over time.

#### \   
##Project Site Map
Insert a text-based overview and site map of the major pages and branch points here.

>>>>

Classifying TA comments has been informative, but doing it by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

My project for the **Faculty Learning Community for R** was to create an automated comment classifier that can sort TA comments from student report into the same categories that we use for hand-coded data. This automated classifier will be used within the larger NSF-funded research project to:

*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
*  Monitor how the key concepts and points that TAs stress in comments on student writing change over time.

#### \   
##Work Plan
The plan for this project is to:

1.  Compile an anonymized testing dataset of hand-coded TA comments from reports. 
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Explore comment data structure, n-gram frequencies, etc., and identify potential elements for feature engineering.
4.  Write a supervised text classification workflow that uses Naive Bayes to assign TA comments to our pre-defined categories. 
5.  Test permutations of analysis parameters to optimize the classifier and establish baseline accuracy.
6.  Apply the optimized NB classifier to the original FULL comment dataset, identify which comments are being classified incorrectly, and look for potential patterns.
7.  Based on features identified in Steps #3 and #6, write a small set of rule- or REGEX pattern-based searches that can identify comments that are more likely to be classified incorrectly.

After completing the FLC project, my goals are to:

8.  Combine the rules/pattern-based pre-screening process with the optimized Naive Bayes classifier to create a vertically  Ensemble Classifier.
9.  Re-validate the Ensemble Classifier against a subset of ~2000 independently hand-coded TA comments extracted from student lab reports for a different semester than the initial test dataset.
10. Begin using the classifier to monitor TA comment patterns in the ongoing NSF-funded project.

Looking further ahead, I hope to use the work presented here as a baseline for evaluating other potential text classification models that are described in the Background.

##Project Site Map
Insert a text-based overview and site map of the major pages and branch points here.

### 2019-03-13 notes.rmd
update

##General/Site-Wide:
  Embedding the bibliographic items  
    Writing the bibtext  
    Embedding using @baseR method, or [@holder].  
  Modify the icons:

  Set menu so landing page is an "Overview"
  Append this code to end of all pages
```
      #### \   
      ***
```

##Index:
*  Decide on bottom page site map strategy, and attach

@@

####\  
888

>>>>

##General/Site-Wide:

+  Write bibliographic entries for missing citations. 
+  Set menu so landing page of each topic is an "Overview"
+  Modify bottom code of all pages. If has citations, make sure the page has:
```
      #### \   
      ***
```
+  If a page has no citations, REMOVE the bottom code for the horizontal rule. It now is in the footer.
+  Beloe is the standard output format for citation tags in my copy of BibDesk. To make the tags work properly, need to search and replace the pre-string.

```\@[Carpenter2001] is output. Change "\@[" to "[@"```

##Index:
*  Decide on bottom page site map strategy, and attach.

@@

####\

### 2019-03-13 techwriting.Rmd
update

Giving effective feedback is a challenging and complex task that needs to be learned through specific training [@Bazerman1994, @Bazerman2006, @Gottschalk2003, @Graves2011, @Kiefer2002, @NCTENWP, @Perelman2009, @Szymanski2014, @Underwood2006]. Effective feedback will be process-oriented rather than focused on finding and correcting errors. Comments will be understandable and unambiguous. Effective feedback keeps the writer in control and encourages them to revise their draft thoughtfully and thoroughly, and keeps copy-editing feedback to a minimum [@Breidenbach2006, @Ruegg2015, @Trupiano2006]. Over-commenting or giving overly prescriptive comments can push a student writer into a passive role. They treat comments as a to-do list that needs to be completed rather than suggestions for reflection and self-assessment [@Anson2001, @CWPA2011, @Harris1979, @Haswell2006a, @Haswell2006b, @Reynolds2009, @Szymanski2015, @Swilky1991, @Underwood2006].

@@

####\  
***

>>>>

Giving effective feedback is a challenging and complex task that needs to be learned through specific training [@Bazerman1994, @Bazerman2006, @Gottschalk2003, @Graves2011, @Kiefer2002, @NCTENWP, @Perelman2009, @Szymanski2014, @Underwood2006]. Effective feedback will be process-oriented rather than focused on finding and correcting errors. Comments will be understandable and unambiguous. Effective feedback keeps the writer in control and encourages them to revise their draft thoughtfully and thoroughly, and keeps copy-editing feedback to a minimum [@Breidenbach2006, @Ruegg2015, @Trupiano2006]. Over-commenting or giving overly prescriptive comments can push a student writer into a passive role. They treat comments as a to-do list that needs to be completed rather than suggestions for reflection and self-assessment [@Anson2001, @CWPA2011, @Harris1979, @Haswell2006a, @Haswell2006b, @Reynolds2009, @Swilky1991, @Underwood2006].

@@

####\  
***
*Work Cited*
